---
title: "TGI-spinal-power"
author: "A.G. Mitchell"
date: '2022-04-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(simr)
```

# Getting data frame and paths
```{r}
datPath <- '/Users/au706616/Documents/Experiments/SPINALTGI/'
setwd(datPath)

df_phs <- read.csv(file.choose())

# need to recode levels for cold_probe to reduce number of levels from 4 to 2
df_phs$cold_location[df_phs$cold_probe == 'distal'] <- 'dist_rostr' 
df_phs$cold_location[df_phs$cold_probe == 'rostral'] <- 'dist_rostr'
df_phs$cold_location[df_phs$cold_probe == 'proximal'] <- 'prox_caud'
df_phs$cold_location[df_phs$cold_probe == 'caudal'] <- 'prox_caud'
df_phs$cold_location <- as.factor(df_phs$cold_location)
```

Create the first model. There is one model for each rating (instead of using rating as a parameter in the model).
First, going to try this on a very simple model where effect of manipulation (TGI vs. control) will be tested on rating of cold thermode, with subject as random effect
```{r}
# Summarise data per rating, per participant
# although summary not needed!
sum_pain <- aggregate(VASburning~ID*manipulation*condition*cold_probe, median, data = df_phs)
sum_warm <- aggregate(VASwarm~ID*manipulation*condition*cold_probe, median, data = df_phs)
sum_cold <- aggregate(VAScold~ID*manipulation*condition*cold_probe, median, data = df_phs)

# first try this on just one, simple model. The effect of manipulation on cold rating, with ID as a random effect
model.cold <- lmer(VAScold ~ manipulation + (1+manipulation|ID), data = df_phs)
summary(model.cold)
```
# View the coefficients for each person (should vary, due to random slope model)
```{r}
coef(model.cold)
```

# Running power analysis on simple model
```{r}
# extract fixed effects from the model
tgi_effect <- fixef(model.cold)["manipulationTGI"]
# need to estimate minimum desired effect from the change in slope - considering values can range from 0-100, it might be simpler to think of these effect sizes as a proportion of 100 instead
# tgi_effect = -17.05 would perhaps leave an expected effect size of .17
# change the effect size to effect size of interest
fixef(model.cold)["manipulationTGI"] <- -17

# then run power analysis
powerSim(model.cold)
```

Now make a second, more complicated model where cold ratings are regressed by both manipulation and dermatome position (condition)
```{r}
model.cold2 <- lmer(VAScold ~ manipulation + condition + (1+manipulation|ID), 
                    data = df_phs)
summary(model.cold2)
```

Finally a third model where all predictors are added - manipulation, condition and cold location
Still just on cold ratings
```{r}
model.cold3 <- lmer(VAScold ~ manipulation + condition + cold_location +
                      (1+manipulation|ID), 
                    data = df_phs)
summary(model.cold3)
```

